---
title: "Lecture1Intro"
author: "Ainsley McLaughlin"
date: "2024-08-27"
output: html_document
---

```{r setup, include=FALSE}

```

## Baysian:
- Bayesian inference:
- Understand process of priors (Prior selection-guesses about model we are fitting)
    - Why diff priors appropriate in diff situations
    - impact of prior choice on model performance
- Understand the process of Bayesian Interence
    - what are conjugate priors
    - how do samplers work and why do we need them
- Bayesian Hierarchical Regression (mixed effect/ multi level models) (this is the     end of content for the final)
    - for clustered data- for data that is not iid
- Overview of bayesian machine learning techniques and why and when
- Overview of more advanced bayesian methods (time permitting)

Course outline:
prob+ inference
bayesian stat
priors conjugates
bayesian inference
samplers, variational interence
prob programming langs- stan, 
bayesian linear regression in brms
bayesian machine learning overview (time permitting)



```{r cars}

```

```{r cars}

```

```{r cars}

```

```{r cars}

```

```{r cars}

```